---
title: "Regular Expressions 3"
author: "Yuyan Yang"
date: "4/16/2022"
output:
  html_document:
    df_print: paged
---

### Load packages and data
```{r warning = FALSE, message = FALSE}
# loading packages
library(tidyverse)

# loading data
regex3 <- read.csv("Data_RegEx3.csv", header = T)
```

### Create New variable year Based on First Four Digits of tourny_id
```{r warning = FALSE, message = FALSE}
# use mutate to create variable year
# and use str_extract() to extract patterns from tourny_id
# store the result to a new df called regex3_1
regex3_1 <- regex3 %>%
  mutate(year = str_extract(tourney_id, pattern = "^\\d{4}"))

# display top 3 and bottom 3 rows
head(regex3_1, 3)
tail(regex3_1, 3)
```

### Separate ‘score’ variable into 5 columns based on ‘score’ results
```{r warning = FALSE, message = FALSE}
# use gsub() to delete patterns "\\(\\d+)" like (4), (5) from score
# use gsub() to delete patterns "RET" from score
# use separate() to separate scores into 5 sets by space
regex3_2 <- regex3_1 %>%
  mutate(newscore = gsub(pattern = "\\(\\d+)", replacement = "", score)) %>%
  mutate(newscore = gsub(pattern = "RET", replacement = "", newscore)) %>%
  separate(newscore, c("Set.1","Set.2","Set.3","Set.4","Set.5"), sep = " ")

# display top 3 and bottom 3 rows
head(regex3_2, 3)
tail(regex3_2, 3)
```

### Create new variable ’sets.played’by counting non-NAs in Set.1 through Set.5
```{r warning = FALSE, message = FALSE}
# sum all the non-NAs in Set.1 through Set.5 and store in a new variable sets.played
# un-select Set.1 to Set.5 because they are not in the 'REGEX3_markdown.pdf'
regex3_3 <- regex3_2 %>%
  mutate(sets.played = as.numeric(!is.na(Set.1)) + 
           as.numeric(!is.na(Set.2)) + 
           as.numeric(!is.na(Set.3)) + 
           as.numeric(!is.na(Set.4)) + 
           as.numeric(!is.na(Set.5))) %>%
  select(!c(Set.1,Set.2,Set.3,Set.4,Set.5))

# display top 3 and bottom 3 rows
head(regex3_3, 3)
tail(regex3_3, 3)
```

### Create Density Plot of Minutes Played by Number of Sets Played
```{r warning = FALSE, message = FALSE}
# use minutes as x-axis and sets.played as color and fill
# change title and labs
# change position and title of legend
regex3_3 %>%
  ggplot(aes(x = minutes, color = factor(sets.played), fill = factor(sets.played))) +
  geom_density() +
  ggtitle("Density Plots of Minutes Played Per Number of Sets Played") +
  xlab('Minutes Played') +
  ylab('Density') +
  theme(legend.position="top") +
  labs(fill="Sets Played",color = "Sets Played")
```

**interpretation**: It looks like 2 sets played is the most common one and 5 sets played is the least common one. Moreover, the density of 2 sets, 3 sets, 4 sets, and 5 sets follows a declining trend. 

### Create Boxplot of Winner’s Age by Tournament Name
```{r warning = FALSE, message = FALSE}
# change Us to US
regex3_4 <- regex3_3 %>%
  mutate(tourney_name = gsub("^Us", "US", tourney_name))

# boxplot of tourney name and winner age
# color by tourney name
# change labels, title, legend title
regex3_4 %>%
  ggplot(aes(x = tourney_name, y = winner_age)) +
  geom_boxplot(aes(color = tourney_name)) +
  ggtitle('Boxplot Age of Winner in Grand Slam Events') +
  xlab('Tournament Name') +
  ylab('Age of Winner') +
  labs(color="Tournament Name")
```

**interpretation**: It looks like that the median of each tournament is the same in term of age of winner and they roughly cover the same interval. 

### Counts of Matches Won by Top 5 IOC Membership per Tournament
```{r warning = FALSE, message = FALSE}
# find out who are the top 5 IOC  per tournament
top5 <- regex3_4 %>%
  group_by(winner_ioc, tourney_name) %>%
  summarise(count = n(), .groups = 'drop') %>%
  group_by(tourney_name) %>%
  arrange(desc(count)) %>%
  mutate(rank = row_number()) %>%
  filter(rank <= 5)

# x-axis is winner_ioc and y is its matches won per tournament
top5 %>%
  ggplot(aes(x = winner_ioc, y = count)) +
  geom_point() +
  facet_grid(tourney_name ~ .) +
  ylab('Tournaments Won') +
  xlab('Winners Olympic Membership') +
  ggtitle('Top 5 Olympic Memberships per Tournament') +
  theme_bw()
```

**interpretation**: It looks like some teams are always the top 5 team in each tournament like ESP, FRA, and GER (they appear in all four facet). FRA has great performance in term of Olympic memberships in all four tournament. 

### Counts of Matches Won by Top 2 IOC Membership per Year
```{r warning = FALSE, message = FALSE}
# find out who are the top 2 IOC per year
top2 <- regex3_4 %>%
  group_by(winner_ioc, year) %>%
  summarise(count = n(), .groups = 'drop') %>%
  group_by(year) %>%
  arrange(desc(count)) %>%
  mutate(rank = row_number()) %>%
  filter(rank <= 2)

#
top2 %>%
  ggplot(aes(x = year, y = count)) +
  geom_point(aes(color = winner_ioc)) +
  xlab('Year') +
  ylab('Tournaments Won') +
  ggtitle('Top 2 Olympic Memberships per Year') +
  labs(color = 'Olympic Membership') +
  theme_bw()
```

**interpretation**: It looks like the tournaments won is decreasing year by year! And this trend has shown in all three teams. It is a very interesting finding because they are the best teams in Olympics (top 2 Olympic Memberships per year), but there is still a declining trend in term of tournaments won.

### link to GitHub
https://github.com/YuyanYang508/Stat-184-Activities/blob/main/regex3/Regex3.Rmd
