---
title: "Regular Expressions 2"
author: "Yuyan Yang"
date: "4/8/2022"
output:
  html_document:
    df_print: paged
---

### Load Packages
```{r message = FALSE, error = FALSE}
library(tidyverse)
```

### Read in Data
```{r message = FALSE, error = FALSE}
# read in .txt using read.delim()
RegEx2 <- read.delim('Data_RegEx2.txt', 
                     fileEncoding = "UTF-8",
                     header = F,
                     quote = "")

# transform file to a vector
RegEx2V <- RegEx2[['V1']]

# transform vector to matrix, then to proper dimension data frame
playerDf <- as.data.frame(matrix(data = RegEx2V, ncol =16, byrow = TRUE))
```

### First Clean
```{r message = FALSE, error = FALSE}
# task 1: Rename original columns, create Player column first name last name

## name vectors
oldnames = c('V1','V2','V3','V4','V5','V6','V7','V8','V9','V10','V11','V12',
             'V13','V14','V15','V16')
newnames = c('Position', 'Player', 'FinalScore', 'THRU',
             'TODAY', 'SGPutt', 'SGARG','SGAPP','SGOTT',
             'SGT2G','SGTOT','DrDist','DrAcc','GIR',
             'FairProx','Scrambling')

## rename original columns
playerDf <- playerDf %>%
  rename_at(vars(oldnames), ~ newnames)

## deal with player's name
playerDf <- playerDf %>% 
  extract(Player, c('Last','First'),"(.+)\\s{1}(.+)") %>%
  unite("Player", First:Last, sep = " ")

# task 2: Remove Position equal to WD (i.e. player withdrew)
playerDf <- playerDf %>%
  filter(!Position == 'WD')

# task 3: Convert hyphens to NAs (view data file, right click hyphened cell, click Inspect, copy that hyphen element to get correct size of hyphen to change to NAs)
playerDf <- playerDf %>%
  mutate(across(everything(), ~gsub("—",NA,.)))

# task 4: Print head and tail to see if results match images provided
head(playerDf)
tail(playerDf)
```

### Second Clean
```{r message = FALSE, error = FALSE}
# task 1: Convert SGTOT and GIR to Numeric while dropping %
playerDf  <- playerDf %>%
  mutate(SGTOT = gsub("%","",SGTOT)) %>%
  mutate(GIR = gsub("%","",GIR)) %>%
  mutate(SGTOT = as.numeric(SGTOT)) %>%
  mutate(GIR = as.numeric(GIR))

# task 2: Create new column Cut based on values of Position where CUT is Missed Cut and Made Cut is otherwise
playerDf$Cut <- ifelse(playerDf$Position == 'CUT', "Missed Cut", "Made Cut")
```

### Create Boxplots
```{r}
# Task 1: Boxplot Shots Gained Total
playerDf %>%
  ggplot(aes(y = SGTOT, color = Cut)) +
  geom_boxplot() +
  ggtitle("Boxplot Shots Gained Total") +
  ylab('Shots Gained Total')
```

It seems like the Missed Cut group has a smaller median than made cut group in term of Shots Gained Total. I will even guess they are significant different based on this box plot.

```{r}
# Task 2: Boxplot Greens Hit in Regulation
playerDf %>%
  ggplot(aes(y = GIR, color = Cut)) +
  geom_boxplot() +
  ggtitle("Boxplot Greens Hit in Regulation") +
  ylab('Greens in Regulation')
```

It seems like, again, missed cut group has a smaller median than made cut group in term of Greens in Regulation. But at least the difference is close this time. 